---
title: "R Insights Workshop - Session 1"
author: "Matthias Gloël"
date: "05 January 2018"
fontsize: 18pt
mainfont: Arial

output: 
  pdf_document:
    latex_engine: xelatex


---

# The most important command
The **help** command opens the documentation 
```{r}
help(read.csv)
```
# Getting and Setting Working Directory

## 1. Check your current working directory:

```{r}
getwd()
```

## 2. Set the working directory to the folder where your data file is located.
**For example "/Users/YourName/Rtutorial/" **

```{r}
setwd("/Users/matthiasgloel/RWorkShop/")
```

## 3. List files in your folder 
```{r, eval=FALSE}
list.files()
```

# Reading in a CSV file from Working Directory
```{r}
data_set <- read.csv("nt-data-example.csv")
```


# Getting some useful information about the data frame
<p>  
Shape of the data frame
```{r}
# Number of columns
ncol(data_set) 
# Number of rows
nrow(data_set)
# Dimenstions
dim(data_set)
```
<p>
The **str** command yields information about the structure of the data frame, i.e. which column contains what type of data (e.g. factors, numbers, text, dates etc. )

```{r}
str(data_set)
```
<p>
The **summary** command yields information about the distribution of the data in the specified column. In order to access a specific column of a data frame you need to type $
<p>
Print first ten rows of data_set
```{r}
head(data_set)
```
<p>
```{r}
summary(data_set$age)
```

# Frequency Tables
```{r}
table(data_set$gender)
```

```{r}
table(data_set$gender, data_set$income)
```

# Aggregating Data (Pivoting)
There are several ways to do that.<br>
The function *tapply* is one of them.
```{r}
## tapply(Summary Variable, Group Variable, Function)
tapply(data_set$age, data_set$city_size, mean)
```